---
title: "Prediction Assignment Writeup"
output: html_document
---

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

# Understanding the Data
We first load the dataset and understanting about the data.
```{r Load Libraries and Data, warning=FALSE, message=FALSE, cache=FALSE}
library(caret)
library(dplyr)
library(RCurl)
library(gridExtra)
library(corrplot)


training <- read.csv("data/pml-training.csv")
testing <- read.csv("data/pml-testing.csv")

dim(training)
dim(testing)
```

## Choosing Predictor Variables
Because of the number of variables (160, the model was taking too long to build and so we decided to use the data that had complete data.  These included belt, arm, dumbbell and forearms)

For more information about the dataset go here >> http://groupware.les.inf.puc-rio.br/work.jsf?p1=11201

Here are the predictor variables selected

```{r GoodDataSearch}
removeMissingData <- sapply(testing, function (x) any(is.na(x) | x == ""))
definingPredictorVariables <- !removeMissingData & grepl("belt|[^(fore)]arm|dumbbell|forearm", names(removeMissingData))
PredictorVariables <- names(removeMissingData)[definingPredictorVariables]
PredictorVariables
```

We have reduced the variables to analyze significantly.
```{r SubsettingNonPredictors}
onlyPredictors <- c("classe", PredictorVariables)
training <- training[, onlyPredictors]
dim(training)
```


# Build the model
The model uses k-mode cross validation and is run 3 times on 10-fold data starting with 70% of the data for training and 30% for validating the model.

Below the accuracy will test the prediction accuracy derived from the training data against the validation data
```{r, cache=TRUE, warning=FALSE}
nuFoldData <- 10
nuRepeats <- 3
testDataSize <- 0.7
base <- read.csv("data/pml-training.csv")
tempTraining <- createDataPartition(base$classe, p = testDataSize, list = FALSE)
training <- base[tempTraining,]
validation <- base[-tempTraining,]

#Using all variables
tempTrControl <- trainControl(method = "repeatedcv", number = nuFoldData, repeats = nuRepeats)
trainedModel <- train(classe ~ ., data = training[,c("classe", "roll_belt","roll_arm", "roll_dumbbell", "roll_forearm", "pitch_belt", "pitch_arm", "pitch_dumbbell", "pitch_forearm", "yaw_belt", "yaw_arm", "yaw_dumbbell", "yaw_forearm")],
                method = "rf", ntree = nuFoldData, trControl = tempTrControl)

confusionMatrix(trainedModel, newdata = predict(trainedModel, newdata = validation))
```

## Predicting events using testing data
Let's test our data using the test data given above
```{r TestingTheModel, warning=FALSE, message=FALSE}
nuPredictions <- 20
finalData <- rep(NA, nuPredictions)
for(i in 1:nuPredictions){
  tmpID <- filter(testing, problem_id == i)
  finalData[i] <- as.vector(predict(trainedModel, newdata = tmpID))
}
finalData
```
